---
title: "Social media dataset"
author: "Krishna Niveditha Pallem"
output: html_document
---
  
#### Loading the Dataset
  
```{r}
library(readr)
library(HSAUR2)
library(SciViews)
library(scatterplot3d)
library(car)
library(lattice)
library(GGally)
library(ggplot2)
library(ggridges)
library(ggvis)
library(ggthemes)
library(cowplot)
library(gapminder)
library(gganimate)
library(dplyr)
library(tidyverse)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(Hotelling)
library(stats)
library(biotools)
library(factoextra)
library(FactoMineR)
library(ggfortify)
library(psych)
library(corrplot)
library(devtools)
library(cluster)
library(magrittr)
library(NbClust)
library(MASS)
library(leaps)
```

```{r}
media <- read.csv("/Users/pallemkrishnaniveditha/Downloads/Socialmediacleaned.csv")
attach( media )
```

**Linear Discriminant Analysis **

```{r}
library(MASS)

media$Tired.waking.up.in.morning <- factor(media$Tired.waking.up.in.morning)

# Apply LDA
media_lda <- lda(Tired.waking.up.in.morning ~ Instagram_Hours + LinkedIn_Hours + Twitter_Hours + 
                 Whatsapp_Wechat_hours + Reddit_hours + Youtube_hours + OTT_hours + 
                 Mood_Productivity + Trouble.falling.asleep, data = media)

# Summary of the LDA model
summary(media_lda)

# Plotting the LDA model
plot(media_lda)

# Optionally, predicting and evaluating the model
media_predictions <- predict(media_lda)
table(predicted = media_predictions$class, actual = media$Tired.waking.up.in.morning)
```

**Inference**

* LDA was used for classifying individuals into groups based on their responses and social media usage. It offered another perspective on how different variables contribute to distinguishing between different classes, such as varying levels of tiredness upon waking.
* Predictions from LDA were evaluated against actual data to assess model accuracy and practical applicability.