---
title: "Social media dataset"
author: "Krishna Niveditha Pallem"
output: html_document
---
  
#### Loading the Dataset
```{r}
library(readr)
library(HSAUR2)
library(SciViews)
library(scatterplot3d)
library(car)
library(lattice)
library(GGally)
library(ggplot2)
library(ggridges)
library(ggvis)
library(ggthemes)
library(cowplot)
library(gapminder)
library(gganimate)
library(dplyr)
library(tidyverse)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(Hotelling)
library(stats)
library(biotools)
library(factoextra)
library(FactoMineR)
library(ggfortify)
library(psych)
library(corrplot)
library(devtools)
library(cluster)
library(magrittr)
library(NbClust)
library(MASS)
library(leaps)
```

```{r}
media <- read.csv("/Users/pallemkrishnaniveditha/Downloads/Socialmediacleaned.csv")
attach( media )
```

  
  ** Logistic regression **

```{r}
# Load necessary libraries
library(dplyr)
table(media$Tired.waking.up.in.morning)
media$Tired.waking.up.in.morning <- ifelse(media$Tired.waking.up.in.morning == "Yes", 1, 0)
media$Tired.waking.up.in.morning <- factor(media$Tired.waking.up.in.morning, levels = c(0, 1))
logit_model <- glm(Tired.waking.up.in.morning ~ Instagram_Hours + LinkedIn_Hours + Twitter_Hours, 
                   data = media, family = binomial)

# Summary of the model to view coefficients and statistics
summary(logit_model)
```

```{r}
library(broom)

# Get a tidy summary of the model
tidy(logit_model)

# Check for multicollinearity
library(car)
vif(logit_model)  # Variance Inflation Factor should ideally be below 5

# Model diagnostics plots
par(mfrow = c(2, 2))
plot(logit_model)
```

```{r}
# Predict probabilities
predicted_probs <- predict(logit_model, type = "response")

# Convert probabilities to binary outcome based on a 0.5 threshold
predicted_class <- ifelse(predicted_probs > 0.5, "Yes", "No")

# Confusion Matrix to evaluate model
table(Predicted = predicted_class, Actual = media$Tired.waking.up.in.morning)

# Calculate accuracy
mean(predicted_class == media$Tired.waking.up.in.morning)
```

**Inference**

* Logistic regression was used to predict binary outcomes such as whether a person felt tired waking up in the morning based on their social media usage. The model provided estimates of how each predictor variable (e.g., hours on Instagram, LinkedIn, Twitter) influences the likelihood of feeling tired.
* The analysis included model diagnostics to check for multicollinearity and model fit, ensuring reliable results.